---
title: "ENVS 193DS Homework 03"
author: "Elena Donelson"
date: today
format: pdf
editor: source
execute:
  message: false
  warning: false
---

# Setup

```{r}
# loading in packages
library(tidyverse)
library(janitor)
library(gt)
library(rstatix)

# reading in kelp data
kelp <- read_csv("temp-kelp.csv")

# reading in my coffee data
coffee_data <- read_csv("coffee_data 2.csv")
```

# Problem 1. Giant kelp fronds

## a.

You could use either a Pearson correlation test or a simple linear regression to determine the strength of the relationship between ocean temperature (°C) and giant kelp frond elongation rate (cm day⁻¹). Both tests analyze the relationship between two continuous quantitative variables, but a Pearson correlation measures the strength and direction of the relationship, while a linear regression tests how changes in ocean temperature predict changes in giant kelp frond elongation rate.

## b.

```{r}
# creating a scatter plot showing the relationship between ocean temperature
# and giant kelp frond elongation rate
ggplot(
  kelp,
  aes(
    x = temp_c,
    y = kelp_elong
  )
) +

  # plotting individual data points
  geom_point(
    color = "darkgreen",
    size = 3
  ) +

  # adding a trend line
  geom_smooth(
    method = "lm",
    se = FALSE,
    color = "steelblue"
  ) +

  # changing the theme
  theme_classic() +

  # labeling the axes and title
  labs(
    title = "Relationship between ocean temperature and giant kelp frond elongation rate",
    x = "Ocean temperature (°C)",
    y = "Giant kelp frond elongation rate (cm day⁻¹)"
  )
```













