---
title: "ENVS 193DS Homework 03"
author: "Elena Donelson"
date: today
format: pdf
editor: source
execute:
  message: false
  warning: false
---

# Setup

```{r}
# loading in packages
library(tidyverse)
library(janitor)
library(gt)
library(rstatix)

# reading in kelp data
kelp <- read_csv("temp-kelp.csv")

# reading in my coffee data
coffee_data <- read_csv("coffee_data 2.csv")
```

# Problem 1. Giant kelp fronds

## a.

You could use either a Pearson correlation test or a simple linear regression to determine the strength of the relationship between ocean temperature (°C) and giant kelp frond elongation rate (cm day⁻¹). Both tests analyze the relationship between two continuous quantitative variables, but a Pearson correlation measures the strength and direction of the relationship, while a linear regression tests how changes in ocean temperature predict changes in giant kelp frond elongation rate.

## b.

```{r}
# creating a scatter plot showing the relationship between ocean temperature
# and giant kelp frond elongation rate
ggplot(
  kelp,
  aes(
    x = temp_c,
    y = kelp_elong
  )
) +

  # plotting individual data points
  geom_point(
    color = "darkgreen",
    size = 3
  ) +

  # adding a trend line
  geom_smooth(
    method = "lm",
    se = FALSE,
    color = "steelblue"
  ) +

  # changing the theme
  theme_classic() +

  # labeling the axes and title
  labs(
    title = "Relationship between ocean temperature and giant kelp frond elongation rate",
    x = "Ocean temperature (°C)",
    y = "Giant kelp frond elongation rate (cm day⁻¹)"
  )
```

## c. 

### Checking assumptions

```{r}
# creating a linear model to test the relationship between ocean temperature
# and giant kelp frond elongation rate
kelp_lm <- lm(
  kelp_elong ~ temp_c,
  data = kelp
)

# creating diagnostic plots to check linear regression assumptions
plot(kelp_lm)
```

I checked the assumptions for a simple linear regression, including linearity, normality of residuals, and equal variance. I checked these assumptions using the diagnostic plots from the linear model. Based on the plots, the assumptions appear reasonable enough to continue with the linear regression test.

### Running the test

```{r}
# running a simple linear regression to test whether ocean temperature
# predicts giant kelp frond elongation rate
summary(kelp_lm)
```

A simple linear regression showed a significant relationship between ocean temperature and giant kelp frond elongation rate (F = 26.93, df = 1, 30, p < 0.001). Temperature was a significant predictor of elongation rate (β = -0.432, p < 0.001), indicating that kelp frond elongation rate decreased as ocean temperature increased. The model explained approximately 47.3% of the variation in elongation rate (R² = 0.473).

## d. 

To evaluate the relationship between ocean temperature and giant kelp frond elongation rate, I used a simple linear regression. I used this test because both variables are continuous quantitative variables, and I wanted to see whether ocean temperature could predict changes in giant kelp frond elongation rate.

The linear regression showed a significant negative relationship between ocean temperature and giant kelp frond elongation rate (F = 26.93, df = 1, 30, p < 0.001, R² = 0.473). As ocean temperature increased, giant kelp frond elongation rate decreased, suggesting that warmer ocean temperatures are associated with slower giant kelp growth.

## e.

The results suggest that higher ocean temperatures are associated with slower giant kelp growth. This could mean that continued ocean warming may negatively affect giant kelp forests by reducing their growth rates and overall health. Since giant kelp provides important habitat for many marine species, monitoring ocean temperatures and kelp growth could help identify areas that may be more vulnerable to climate change.







